---
title: "206_Assignment_4_Lobsters"
author: "Lydia Bleifuss and Camila Bobroff"
date: "11/15/18"
output:
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Analyzing California Spiny Lobster Abundance, Size and Fishing Implications

###A brief assessment conducted with Spiny Lobster samples taken from five sites in the Santa Barbara Channel from 2012 through 2017

#####Lydia Bleifuss & Camila Bobroff
#####Bren School of Environmental Science and Management
#####University of California, Santa Barbara 
```{r, include = FALSE}
library(tidyverse)
library(pwr)
library(knitr)
library(plotly)
library(extrafont)
library(ggrepel)
library(kableExtra)
library(effsize)
library(dplyr)
library(vcdExtra)
library(car)
library(ggpubr)
library(ggplot2)
library(ggsignif)
library(xtable)
```

```{r, include = FALSE}
lob_size_abun <- read_csv("lobster_size_abundance.csv")
lob_trap <- read_csv("lobster_traps.csv")
```
####Introduction

Southern California’s Spiny Lobster has been fished commercially since the 1800s. The Santa Barbara Channel Spiny Lobster fishery is one of California’s most economically critical, and has also undergone concerning population changes within the last decade.2 Spiny Lobster were being harvested at record rates from the late 2000s to 2010. Improvements in fishing technology and increases in recreational fishing however, combined with an observed 15% decline in reported lobsters caught, necessitated improved monitoring.2 

The Santa Barbara Coastal Long Term Ecological Research Project (SBC LTER), operating within University of California, Santa Barbara, has lead data collection surrounding this topic in conjunction with the National Science Foundation’s Long Term Ecological Research Network since the 1980s. In 2012, five Santa Barbara Coastal sites which were established close to the mainland: Arroyo Quemado (AQUE), Naples Reef (NAPL), Mohawk Reef (MOHK), Isla Vista (IVEE), and Carpinteria (CARP). Of those, Naples and Isla Vista were designated no-take marine protected areas (MPAs). MPAs, or oceanic corridors defined by boundaries in which human activity is limited and fishing is restricted, are established to promote marine biodiversity. In some cases, MPA parameters are selected with specific species targets in mind.1 

Although, Spiny Lobster size, abundance, and fishing pressure in the Santa Barbara Chanel has been monitored since 2012, overall lack of previously collected data for this fishery confines analysis relative to recent years.3 In this report, we utilize data that has been collected by the SBC LTER project from 2012 through 2017 at the five sites. The purpose of this analysis is to describe lobster abundance and size relative to trapping presence and to determine the significance of MPA designation in relation to lobster size and abundance over time. 

####Data Analysis and Statistical Methods 

Our data was collected by dive teams from 2012 through 2017 and summarized by Daniel Reed via SBC.3 The dataset produced two tables, the first indicating lobster abundance and size (carapace length in mm) collected at the five LTER sites annually before the start of the fishing season in late summer. The second displays the number of commercial trap floats in each site, indicating fishing pressure. Every two to four weeks throughout the October to March fishing season, total trap counts are collected at the five LTER sites and four additional sites near the mainland. Eight of the nine sites are LTER. For the purposes of our analysis, we will only be utilizing the five LTER sights included across both tables. 

The information compiled in this dataset including date, site, site location and MPA designation, lobster length (mm) and relative count, and total trap counts per site, was utilized to calculate mean sample statistics. We used this data to detect any associations between lobster abundance and fishing pressure throughout the data collection years, compare mean lobster size across the 5 sites in 2017, changes in mean size at MPA and non-MPA sites between 2012 and 2017, and the proportion of legal lobsters at all 5 sites in 2017. More specifically, we used a one-way ANOVA with a post-hoc Tukey’s HSD test to compare mean lobster sizes, and used both one-sided and two-sided two sample Student’s T-tests (α = 0.025 and α = 0.05 respectively) to determine significance or lack thereof for changes in lobster size. Our analysis, including graphs and tables, was produced using R Statistical Software through R Studio (v. 1.1456).

###Results and Discussion

*Lobster abundance and fishing pressure (2012 - 2017)*
To gain an initial impression of lobster abundance and fishing pressure we compared all five sites, MPA and non-MPA, contrasting relative abundance and trap total data. No traps were observed at the two MPA sites, Isla Vista and Naples Reef, from 2012 to 2017. Overall increases in Isla Vista and Carpinteria lobster abundance is notable, and dramatically so from 2016 to 2017. However, while Carpinteria trap totals decreased significantly from 2016-2017, Isla Vista totals remained constant at zero. Despite 2015 count peaks for Mohawk and Naples, abundance levels at Mohawk Reef and Naples Reef lobster in 2017 are relatively lower than Isla Vista and Carpinteria. All three non-MPA sites appear to decrease trap totals from 2016 to 2017. Despite Isla Vista and Naples MPA designation, abundance at these two sites does not appear to react similarly year to year.

```{r, echo = FALSE, message = FALSE}
#Producing mean abundance by year
abun_year <- lob_size_abun %>%
  group_by(YEAR, SITE) %>% 
  filter(COUNT != "0") %>% 
  summarize(
    sum_abun = round(sum(COUNT), digits = 2)
  ) %>% 
  rename(Site = SITE)
#Producing mean trap count by year

trap_year <- lob_trap %>% 
  filter( SITE != "AHND", SITE != "ABUR", SITE != "GOLB", SITE != "AHND to AQUE") %>% 
  group_by(YEAR, SITE) %>%
  summarize(
    sum_trap = round(sum(TRAPS), digits = 2)
  ) %>% 
  rename(Site = SITE)

#Join the data sets together so they pull the same year and corresponding data 

abun_trap <- full_join(abun_year, trap_year) %>% 
  select(YEAR, Site, sum_abun, sum_trap)
```

```{r, echo = FALSE, message = FALSE}
#Creating Scatter-plots for question 1. 

#Abundance Totals 
abun_point <- ggplot(abun_trap, (aes(x = YEAR,
      y = sum_abun,label = ""))) +
  geom_point(aes(color = Site)) +
  theme_classic() +
  scale_size(range = c(0,9)) +
  geom_line(aes(color = Site), size = 0.5) +
   scale_y_continuous(expand = c(0,0), limits = c(0,800)) +
  labs(x = "Year", y = "Total Spiny Lobster Abundance") +
  theme(text = element_text(family = "Times New Roman"))+
  scale_colour_discrete(name="Site",
                         breaks=c("AQUE", "CARP", "IVEE", "MOHK", "NAPL"),
                         labels=c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk Reef", "Naples Reef")) #Change legend labels

abun_point
```

**Figure 1. Lobster Abundance in the Santa Barbara Channel (2012 - 2017).**  
Counts of California spiny lobster collected annually in late summer before the fishing season. Data used with permission from SBC LTER^3^.
```{r echo = FALSE, message = FALSE}
#Trap Totals
trap_point <- ggplot(abun_trap, (aes(x = YEAR,
      y = sum_trap,label = ""))) +
  geom_point(aes(color = Site))+
  theme_classic() +
  scale_size(range = c(0,9)) +
  geom_line(aes(color = Site), size = 0.5) +
  scale_y_continuous(expand = c(0,0), limits = c(0,1300)) +
  labs(x = "Year", y = "Site Trap Totals") +
  theme(text = element_text(family = "Times New Roman")) +
  scale_colour_discrete(name="Site",
                         breaks=c("AQUE", "CARP", "IVEE", "MOHK", "NAPL"),
                         labels=c("Arroyo Quemado", "Carpinteria", "Isla Vista", "Mohawk Reef", "Naples Reef")) #Change legend labels

trap_point
```

**Figure 2. Lobster Traps in the Santa Barbara Channel (2012 - 2017).**  
Counts of commercial trap floats collected every two to four weeks during the fishing season. Data used with permission from SBC LTER^3^.

*Compare mean lobster size by site in 2017*
Lobster carapace length (mm) differed significantly between the five LTER sites in 2017 (one-way ANOVA, F(4, 1663) = 3.42, p = .009; Table 2). Post-hoc analysis by Tukey’s HSD revealed that mean lobster carapace length differs significantly between Naples Reef and Carpinteria and between Naples Reef and Isla Vista (pairwise p = 0.023 and 0.004, respectively). While lobsters at one of the MPA sites (Naples Reef) did have larger carapace length than those at a non-MPA site (Carpinteria), lobsters at Isla Vista, also protected, did not show significantly different results from a non-MPA site. These results do not indicate that MPA designation is associated with lobster size. 

```{r, echo = FALSE, message = FALSE}

#Step 1: Get data into case/tidy format

count <- as.data.frame(lob_size_abun) #Coerce to data.frame so we can use expand.dft()
lob_size_tidy <- expand.dft(count, freq = "COUNT")#Expand data so it is in tidy format

lob_size_tidy2017 <- lob_size_tidy %>% 
  filter(YEAR == 2017) %>%  #Keep observations only from 2017
  mutate(
    mean_size2017 = mean(SIZE),
    sd_size = sd(SIZE),
    sample_size = length(SITE)
  )

#Step 2: Calculate mean lobster sizes (carapace length (mm)) across the five sites for observations in 2017

lob_size_summary <- lob_size_tidy2017 %>% 
  group_by(SITE) %>% 
  summarize(
    mean_size2017 = round(mean(SIZE),2),
    sd_size = round(sd(SIZE),2),
    sample_size = length(SITE)
  )
#Step 3: Test assumptions, exploratory graphs, summary statistics, Levene's test for equal variances

lobster_hists <- ggplot(lob_size_tidy2017, aes(x = SIZE))+
  geom_histogram(aes(fill = SITE)) + 
  facet_wrap(~ SITE)

lobster_qqplot <- ggplot(lob_size_tidy2017, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ SITE)
#Data seem to be normally distributed based on our histograms and qqplots!

# Levene's test for equal variances (> 2 groups)

# H0: Variances are equal
# HA: Variances are unequal

lobster_levene <- leveneTest(SIZE ~ SITE, data = lob_size_tidy2017)
# We reject the null hypothesis of equal variances (p < 0.001)...

variances <- lob_size_tidy2017 %>% 
  group_by(SITE) %>% 
  summarize(
    variance = var(SIZE)
  )
# Because the largest variance is less than 4 times greater than the smallest variances, we can use an ANOVA to compare means.

#Step 4: One-way ANOVA
#Single factor: SITE | Number of levels: 5 (AQUE, CARP, IVEE, MOHK, NAPL) | Random variable: SIZE (carapace length (mm))

#QUESTION: Is there a significant difference in mean carapace length (mm) in AQUE, CARP, IVEE, MOHK, and NAPL?

# H0: Mean lobster carapace length across all sites are equal 
# HA: At LEAST two means differ significantly

lobster_aov <- aov(SIZE ~ SITE, data = lob_size_tidy2017)
# At least two samples were taken from populations with different means (P = .009). Which ones are different?

#Step 5: Post-hoc test (Tukey's HSD)

lobster_ph <- TukeyHSD(lobster_aov)
```

```{r, r ANOVA_table, echo = FALSE, results = 'asis', message = FALSE}
#Step 6: Create table and column graph with error bars indicating ±1 standard deviation

lob_boxplot <- ggplot(lob_size_tidy2017, aes(x = SITE, y = SIZE))+
  geom_boxplot(aes(fill = SITE, colour = SITE), alpha = 0.5, show.legend = FALSE)+
  labs(title="Lobster Sizes by Location in 2017", 
                    x ="Location",
                    y = "Carapace Length (mm)")+
  theme_classic()+
  theme(legend.position = "none")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  stat_compare_means(method = "anova", label.y = 170)+ # Add global p-value
  stat_compare_means(label = "p.signif", method = "t.test",
                     ref.group = "0.5")+ # Pairwise comparison against
    scale_x_discrete(labels=c("AQUE" = "Arroyo Quemado",
                   "CARP" = "Carpinteria",
                   "IVEE" = "Isla Vista",
                   "MOHK" = "Mohawk Reef",
                   "NAPL" = "Naples Reef"))
lob_boxplot

lob_size_graph <- lob_size_summary %>% 
   ggplot(aes(x = SITE, y = mean_size2017)) +
   geom_col(fill = "tomato3", colour = "black", aes(fill = SITE), position = "dodge", show.legend = FALSE) +
  labs(y = "Mean Lobster Carapace Length (mm)", x = "Site")+
  geom_signif(comparisons = list(c("NAPL", "CARP")), annotations = "p = 0.023", y_position = 95, tip_length = 0.1, size = 0.5, textsize = 3)+
   theme_classic()+
  geom_signif(comparisons = list(c("NAPL", "IVEE")), annotations = "p = 0.004", y_position = 110, tip_length = 0.1, size = 0.5, textsize = 3)+
  scale_y_continuous(expand = c(0,0), limits=c(0,130))+
  scale_y_continuous(expand = c(0,0), limits=c(0,130))+
  geom_errorbar(aes(ymin=mean_size2017-sd_size, ymax=mean_size2017+sd_size), width=.2)+
  scale_x_discrete(labels=c("AQUE" = "Arroyo Quemado",
                   "CARP" = "Carpinteria",
                   "IVEE" = "Isla Vista",
                   "MOHK" = "Mohawk Reef",
                   "NAPL" = "Naples Reef"))
lob_size_graph

lob_table2017 <- lob_size_summary %>% 
    mutate(Site=
    case_when(
      SITE == "NAPL" ~ "Naples Reef",
      SITE == "AQUE" ~ "Arroyo Quemado",
      SITE == "MOHK" ~ "Mohawk Reef",
      SITE == "IVEE" ~ "Isla Vista",
      SITE == "CARP" ~ "Carpinteria"
    )) %>% 
  select(Site, mean_size2017, sd_size, sample_size) %>% 
  kable(col.names = c("Site", "Mean Carapace Length (mm)", "Standard Deviation (mm)", "Sample Size"), align = "c") %>% 
  kable_styling(bootstrap_options = c("striped"))
  
lob_table2017
```
**Table 1. Mean lobster size summary by site (2017).**  
Data source: SBC LTER^3^. 
* Indicates an MPA site

**Table 2. Mean lobster size ANOVA results (2017).**
Data source: SBC LTER^3^.

**Figure 3. Mean Lobster Size in the Santa Barbara Channel (2017)**  
2017 mean lobster carapace length (mm) by site, measured in late summer before the fishing season. Error bars indicate ± 1 standard deviation. Brackets indicate *p*-values for significantly different means as determined by one-way ANOVA (F(4, 1663) = 3.4, *p* = 0.009, $\alpha$ = 0.05) with post-hoc Tukey's HSD ($\alpha$ = 0.05). Data used with permission from SBC LTER^3^.


*Changes in lobster size at MPA and non-MPA sites (comparing only 2012 and 2017 sizes)*

```{r, echo = FALSE, message = FALSE}
#Step 1: Calculate mean size at each site for 2012 and 2017. Data exploration.

graph_size_change <- lob_size_tidy %>% 
  filter(YEAR == "2012" | YEAR == "2017") %>%
  group_by(YEAR) %>% 
  select(YEAR, SITE, SIZE) %>% 
  ggplot(aes(x = SITE, y = SIZE, group = YEAR))+
    geom_col(position = 'dodge', aes(fill = YEAR))+
  theme(legend.position = c(2012, 2017))+
  theme_classic()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  scale_y_continuous(expand = c(0,0), limits=c(0,130))+
  scale_y_continuous(expand = c(0,0), limits=c(0,130))+
  scale_x_discrete(labels=c("AQUE" = "Arroyo Quemado",
                   "CARP" = "Carpinteria",
                   "IVEE" = "Isla Vista",
                   "MOHK" = "Mohawk Reef",
                   "NAPL" = "Naples Reef"))
graph_size_change

lob_size_IVEE <- lob_size_tidy %>% 
  filter(SITE == "IVEE") %>%  
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  group_by(YEAR)

IVEE_summary <- lob_size_IVEE %>% 
  summarize(
    mean_size = mean(SIZE),
    sample_size = length(SIZE), #Sample sizes are VERY different, only 26 observations for IVEE in 2012
    sd_size = sd(SIZE)
  )

lob_size_NAPL <- lob_size_tidy %>% 
  filter(SITE == "NAPL") %>%  
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  group_by(YEAR)

NAPL_summary <- lob_size_NAPL %>% 
  summarize(
    mean_size = mean(SIZE),
    sample_size = length(SIZE) #Sample sizes are very different again, only 6 observations in 2012 for NAPL! Might be something to mention in the write up..
  )

lob_size_AQUE <- lob_size_tidy %>% 
  filter(SITE == "AQUE") %>%  
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  group_by(YEAR)

AQUE_summary <- lob_size_AQUE %>% 
  summarize(
    mean_size = mean(SIZE),
    sample_size = length(SIZE)
  )

lob_size_MOHK <- lob_size_tidy %>% 
  filter(SITE == "MOHK") %>%  
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  group_by(YEAR)

MOHK_summary <- lob_size_MOHK %>% 
  summarize(
    mean_size = mean(SIZE),
    sample_size = length(SIZE),
    sd_size = sd(SIZE)
  )

lob_size_CARP <- lob_size_tidy %>% 
  filter(SITE == "CARP") %>%  
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  group_by(YEAR)

CARP_summary <- lob_size_CARP %>% 
  summarize(
    mean_size = mean(SIZE),
    sample_size = length(SIZE)
  )

#Step 2: Test assumptions (each site for 2012 and 2017)
hists2017 <- ggplot(lob_size_tidy2017, aes(x = SIZE))+
  geom_histogram(aes(fill = SITE))+
  facet_wrap(~SITE)

qqplots2017 <- ggplot(lob_size_tidy2017, aes(sample = SIZE)) +
  geom_qq()+
  facet_wrap(~SITE)
#Based on our histograms and qqplots, our 2017 data for each site looks normal

lob_size_tidy2012 <- lob_size_tidy %>%
  filter(YEAR == "2012")

hists2012 <- ggplot(lob_size_tidy2012, aes(x = SIZE))+
  geom_histogram(aes(fill = SITE))+
  facet_wrap(~SITE)

qqplots2012 <- ggplot(lob_size_tidy2012, aes(sample = SIZE)) +
  geom_qq()+
  facet_wrap(~SITE)
#Based on our histograms and qqplots, our 2012 data for each site look normal enough... IVEE is the farthest from normal, with only 6 observations. Might be worth noting in the write-up.

# F-test for equal variances
# H0: Variances are equal (ratio of variances = 1)
# HA: Variances are not equal (ratio of variances is NOT = 1)

IVEE_ftest <- lob_size_IVEE %>% 
  var.test(SIZE ~ YEAR, data = .) # data = . is saying to access the most recent datafarme used

IVEE_ftest # Retain the null hypothesis of equal variances (p-value = 0.30)
# We can override the default setting in t.test() function of var.equal = FALSE

NAPL_ftest <- lob_size_NAPL %>% 
  var.test(SIZE ~ YEAR, data = .) # data = . is saying to access the most recent datafarme used

# Retain the null hypothesis of equal variances (p-value = 0.76)
# We can override the default setting in t.test() function of var.equal = FALSE

AQUE_ftest <- lob_size_AQUE %>% 
  var.test(SIZE ~ YEAR, data = .) # data = . is saying to access the most recent datafarme used

# Retain the null hypothesis of equal variances (p-value = 0.30)
# We can override the default setting in t.test() function of var.equal = FALSE

MOHK_ftest <- lob_size_MOHK %>% 
  var.test(SIZE ~ YEAR, data = .) # data = . is saying to access the most recent datafarme used

# Retain the null hypothesis of equal variances (p-value = 0.15)
# We can override the default setting in t.test() function of var.equal = FALSE

CARP_ftest <- lob_size_CARP %>% 
  var.test(SIZE ~ YEAR, data = .) # data = . is saying to access the most recent datafarme used

# Retain the null hypothesis of equal variances (p-value = 0.20)
# We can override the default setting in t.test() function of var.equal = FALSE

#Step 3: One sided, two sample t-test to compare lobster size at MPA sites (IVEE and NAPL)

#HO: Lobsters are NOT significantly larger in 2017 than in 2012
#HA: Lobsters are significantly larger in 2017 than in 2012

IVEE_ttest <- lob_size_IVEE %>%
  t.test(SIZE ~ YEAR, var.equal = TRUE, alternative = "less", data = ., conf.level = 0.975)
#P=0.03, reject the null

#Power analysis
IVEE_2012 <- lob_size_IVEE %>% 
 filter(YEAR == "2012") %>% 
 pull(SIZE)
 
IVEE_2017 <- lob_size_IVEE %>% 
 filter(YEAR == "2017") %>% 
 pull(SIZE)

IVEE_effsize <- cohen.d(IVEE_2012,IVEE_2017)
#Small effect size (and remember that IVEE had a VERY small sample size in 2012)

NAPL_ttest <- lob_size_NAPL %>%
  t.test(SIZE ~ YEAR, var.equal = TRUE, alternative = "less", data = ., conf.level = 0.975)
#P=0.25, retain the null

#Power analysis
NAPL_2012 <- lob_size_NAPL %>% 
 filter(YEAR == "2012") %>% 
 pull(SIZE)
 
NAPL_2017 <- lob_size_NAPL %>% 
 filter(YEAR == "2017") %>% 
 pull(SIZE)

NAPL_effsize <- cohen.d(NAPL_2012,NAPL_2017)
#Small effect size

#Step 3: Two sided, two sample t-test to compare lobster size at non-MPA sites (AQUE, MOHK, and CARP)

#HO: Lobster size does NOT significantly different between 2012 and 2017
#HA: Lobster size is significantly different between 2012 and 2017

AQUE_ttest <- lob_size_AQUE %>%
  t.test(SIZE ~ YEAR, var.equal = TRUE, data = .)
#P=0.21, retain the null

#Power analysis
AQUE_2012 <- lob_size_AQUE %>% 
 filter(YEAR == "2012") %>% 
 pull(SIZE)
 
AQUE_2017 <- lob_size_AQUE %>% 
 filter(YEAR == "2017") %>% 
 pull(SIZE)

AQUE_effsize <- cohen.d(AQUE_2012,AQUE_2017)
#Small effect size

MOHK_ttest <- lob_size_MOHK %>%
  t.test(SIZE ~ YEAR, var.equal = TRUE, data = .)

#P<.001, reject the null

MOHK_2012 <- lob_size_MOHK %>% 
 filter(YEAR == "2012") %>% 
 pull(SIZE)
 
MOHK_2017 <- lob_size_MOHK %>% 
 filter(YEAR == "2017") %>% 
 pull(SIZE)

MOHK_effsize <- cohen.d(MOHK_2012,MOHK_2017)
#Medium effect size! (Cohen's d = 0.54)

CARP_ttest <- lob_size_CARP %>%
  t.test(SIZE ~ YEAR, var.equal = TRUE, data = .)

#P=0.182, retain the null
CARP_2012 <- lob_size_CARP %>% 
 filter(YEAR == "2012") %>% 
 pull(SIZE)
 
CARP_2017 <- lob_size_CARP %>% 
 filter(YEAR == "2017") %>% 
 pull(SIZE)

CARP_effsize <- cohen.d(CARP_2012,CARP_2017)
#Negligible effect size

```
Mean lobster carapace lengths at Isla Vista, an MPA site, were significantly larger in 2017 than in 2012 [t(`r IVEE_ttest$parameter`) = `r round(IVEE_ttest$statistic,2)`, *p* = `r round(IVEE_ttest$p.value,3)`, $\alpha$ = 0.025]. Despite these significant findings there is a small effect size (Cohen's d = 0.38) and the number of observations between the two years were very different ($\overline{X}$ = 66.08 ± 12.09 [mm], n = 26, $\overline{X}$ = 71.45 ± 14.32 [mm], n = 606, for 2012 and 2017, respectively). Mean lobster lengths at Mohawk Reef, a non-MPA site, differed significantly between 2012 and 2017 [t(`r MOHK_ttest$parameter`) = `r round(MOHK_ttest$statistic,2)`, *p* = `r round(MOHK_ttest$p.value,3)`, $\alpha$ = 0.05]. Furthermore, there is a medium effect size (Cohen's d = 0.54) between the mean lobster lengths at Mohawk Reef in 2012 and 2017 ($\overline{X}$ = 77.25 ± 10.59 [mm], n = 83, $\overline{X}$ = 72.00 ± 9.28 [mm], n = 178, for 2012 and 2017, respectively).

These results indicate that lobsters at Mohawk Reef, non-MPA, were significantly smaller in 2017 than in 2012, with a medium effect size. While the data show an overall increase in lobster lengths at MPA sites, the magnitude of this increase is small. However, the decrease in lobster size in non-MPA sites, considering the medium effect size is notable. In this case, non-MPA status appears to implicate decrease in lobster length from 2012 to 2017. 


**Table 3. Lobster size measurements (mm) at Isla Vista and Mohawk Reef sites in 2012 to 2017.**  
Values represente mean ± standard deviation, and sample sizes are listed directly below for each site.These comparisons indicate the significant results from the T-test indicated above. Data source: SBC LTER^3^.

Part 4. Proportions of "legal" lobsters at the 5 sites in 2017

The legal minimum carapace size for lobster is 82.6 mm. What proportion of observed lobsters at each site are above the legal minimum? Does that proportion differ significantly across the 5 sites? 
```{r, echo = FALSE, message = FALSE}
#Propper data frame 

lob_chi <- lob_size_tidy2017 %>% 
  mutate(
    limit = case_when(SIZE > 82.6 ~ "legal", SIZE < 82.6 ~ "illegal")
  )

chi_size <- lob_chi %>%
  count(SITE, limit) %>% 
  spread(limit, n) %>% 
  select(-SITE)

rownames(chi_size) <- c("AQUE","CARP","IVEE","MOHK","NAPL")

#Actual proportions:
size_prop <- prop.table(as.matrix(chi_size), 1)
#Chi-squared 
#H0: No significant difference between proportions at the different sites

size_x2 <- chisq.test(chi_size)
```
**Table 4. Counts and proportions of lobsters below and above the legal limit, by site.**
Values indicated are counts, with proportion (by site) in italics. Lobsters above or below the legal limit differ significantly by site ($\chi^2${`r size_x2$parameter`} = 18.50, *p* = 0.001). Data source: SBC LTER)^3^.

There is a significant association between observed lobsters above or below the legal limit (82.6 mm) and location ($\chi^2${`r size_x2$parameter`} = 18.50, *p* = 0.001). Notably, lobsters above the legal limit at Naples Reef, an MPA site, (33%) was remarkably higher than those observed at Mohawk Reef, a non-MPA site (13%). 
